Active Learning for SAT Solver Benchmarking

### 2022-10-10 paper/paper.tex
section2.3
section3.2

To determine solver ranks, we define the adapted scoring function $s : A \rightarrow [1, 2 \cdot k]$ as follows.
\begin{align*}
  \operatorname{s}(a) &:= \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
  &&
  \gamma'_k(a, e) &:= \begin{cases}
  	2 \cdot \gamma_k(a, e) 	& \text{if } \gamma_k(a, e) = k\\
	\gamma_k(a, e)	& \text{otherwise}
  \end{cases}
  \label{eq:rankingeq}
\end{align*}

Discretization would not provide much value if the resulting cluster labels were not predictive of the PAR-2 ranking we seek to predict.
For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}).
This shows that monotonicity and score distances between ranks align quite well between both ranking methods.

% In that sense, we also experiment with a combination of runtime-label ranking and the PAR-2 ranking of the subsample if the difference in label ranking scores drops below a certain threshold.
% Besides that, we additionally consider using the past $l$ model predictions in a voting approach for stability.

That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
This means that we do not lose much discriminatory power using discrete runtime labels.
The predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification, show the clear advantage of runtime discretization.

>>>>

To determine solver ranks, we define the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows.
\begin{align*}
  s_k(a) := \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
  &&
  \gamma'_k(a, e) := \begin{cases}
  	2 \cdot \gamma_k(a, e) 	& \text{if } \gamma_k(a, e) = k\\
	\gamma_k(a, e)	& \text{otherwise}
  \end{cases}
  \label{eq:rankingeq}
\end{align*}

The scoring function $s_k$ induces a ranking among solvers. 
In preliminary experiments, we compared the ranking induced by $s_3$ to the ranking induced by the PAR-2 score which is commonly used in SAT competitions~\cite{FroleyksHIJS21}. 
The Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988} which is very close to the optimal value of 1~\cite{de2016comparing}.

That discretized runtimes are sufficient for discriminating solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different. 
In preliminary experiments, we used a Wilcoxon signed rank test with $\alpha = 0.05$ to compare solver pairs by their runtime vectors.
Within the SAT~Competition~2022 Anniversary track, about \SI{89.95}{\%} of solver pairs are significantly different according to that measure.
The strongly discretized runtimes $\gamma_2$ are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, while $\gamma_3$ is sufficient to distinguish about \SI{87.83}{\%} of solver pairs.

% MI: Enough justification, commented out the following
%For the SAT Competition~2022 Anniversary track instances, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
%Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.

This shows that monotonicity and score distances between ranks align quite well between both ranking methods.
The correlation between $s_k$-induced and PAR-2 induced ranks, the discriminatory power of $\gamma_k$, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
